---
title: "HW3"
output: github_document
---

```{r setup, message=FALSE}
library(tidyverse)
library(readxl)
library(lubridate)
library(knitr)
```

```{r loading data, message=FALSE}
library(p8105.datasets)
data("instacart")
```

The instacart dataset includes data from online grocery orders. The dataset includes `r nrow(instacart)` rows/observations and `r ncol(instacart)` columns/variables. Key variables include identifiers for each order, product, customer, aisle, and department; the day/hour of purchase; and product, aisle, and department names. For example, we can see that `r pull(instacart, product_name)[10]` was ordered from the aisle containing `r pull(instacart, aisle)[10]` in the department including `r pull(instacart, department)[10]` on day `r pull(instacart, order_dow)[10]` of the week.



```{r aisles, message=FALSE, warning=FALSE}
aisles = instacart %>% 
  group_by(aisle, aisle_id) %>%
  tally() %>% 
  arrange(desc(n))
aisles
```

The previous code chunk shows the aisles and aisle IDs in descending order of number of items ordered. There are `r nrow(aisles)` aisles, and the aisles with the most items ordered are `r pull(aisles, aisle)[1]` with `r pull(aisles, n)[1]` orders, `r pull(aisles, aisle)[2]` with `r pull(aisles, n)[2]` orders, and `r pull(aisles, aisle)[3]` with `r pull(aisles, n)[3]` orders.

Limited to aisles with more than 10000 items observed, the plot generated below shows the number of items ordered in each aisle.

```{r aisles cleaning and plot, message=FALSE}
instacart %>%
  group_by(aisle) %>% 
  tally() %>% #creates n variable so that I can make aisle a factor variable with levels based on n
  ungroup() %>% 
  mutate(aisle = factor(aisle, levels = aisle[order(n)])) %>% #ordered to make plot from most to least items ordered
  filter(n > 10000) %>% 
  ggplot(aes(x = aisle, y = n)) +
  geom_col() +
  coord_flip() +
  labs(title = "Items Ordered per Aisle",
       x = "Aisle",
       y = "Count of Items Ordered"
       ) +
  theme(text = element_text(size = 10))
```



The table below shows the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits” along with the number of times each item was ordered.

```{r table with popular item counts}
popular = instacart %>% 
  filter(aisle == "baking ingredients" | aisle == "dog food care" | aisle == "packaged vegetables fruits") %>%
  group_by(aisle, product_name) %>% 
  summarize(count = n()) %>% 
  top_n(3, count) %>% 
  arrange(desc(count)) %>% 
  rename("Aisle" = aisle, "Product Name" = product_name, "Item Counts" = count)

kable(popular, format = "html", caption = "Most Popular Items per Aisle")
```


The 2x7 table below shows the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week.

```{r mean hour of order table}
apples_icecream = instacart %>%
  filter(product_name == "Pink Lady Apple" | product_name == "Coffee Ice Cream") %>%
  group_by(product_name, order_dow) %>% 
  summarize(meanhr = mean(order_hour_of_day)) %>%
  mutate(order_dow = order_dow + 1,
         order_dow = wday(order_dow, label = TRUE)) %>% #recode numbers to days of week
  spread(key = order_dow, value = meanhr) %>% 
  select(product_name, Sun, Mon, Tue, Wed, Thu, Fri, Sat) %>% 
  rename("Product Name" = product_name) #to make the column name more presentable

kable(apples_icecream, format = "html", caption = "Mean Hour at which Pink Lady Apples and Coffee are Ordered", digits = 2)
```


#PROBLEM 2

The BRFSS dataset can be accessed from the p8105 library.
```{r loaded brfss, message=FALSE}
data("brfss_smart2010")
```

The data is cleaned below to use appropriate variable names, to focus on the “Overall Health” topic, to include only responses from “Excellent” to “Poor”, and to organize responses as a factor taking levels ordered from “Poor” to “Excellent”.

```{r cleaning brfss, message=FALSE}
brfss = brfss_smart2010 %>% 
  janitor::clean_names() %>% 
  mutate(response = factor(response, levels = c("Poor", "Fair", "Good", "Very good", "Excellent"))) %>% #selects just those between poor and excellent
  filter(topic == "Overall Health")
```


The tables below show states observed at 7 or more locations in 2002 vs 2010.
```{r}
states_2002 = brfss %>%
  filter(year == 2002) %>% 
  group_by(locationabbr) %>% #locationabbr lists by state
  distinct(locationdesc) %>% #locationdesc describes states by counties
  tally() %>%
  filter(n >= 7)

states_2010 = brfss %>%
  filter(year == 2010) %>% 
  group_by(locationabbr) %>% #locationabbr lists by state
  distinct(locationdesc) %>% #locationdesc describes states by counties
  tally() %>%
  filter(n >= 7)

kable(states_2002, format = "html", caption = "2002")
kable(states_2010, format = "html", caption = "2010")

```
Above, we can see that there were `r nrow(states_2002)` states that were observed at 7 or more locations in 2002: Connecticut, Florida, Massachussetts, North Carolina, New Jersey, and Pennsylvania.
Likewise, in 2010, there were `r nrow(states_2010)` states observed at 7 or more location in 2010: California, Colorado, Florida, Massachussetts, Maryland, North Carolina, Nebraska, New Jersey, New York, Ohio, Pennsylvania, South Carolina, Texas, Washington. Based on this, we can tell that observations expanded from 2002 to 2010.


The dataset below shows year, state, and an average data value across locations within a state, limited to Excellent responses. From this dataset, a spaghetti plot was generated to depict the average value over time within each state.

```{r excellent dataset and spaghetti plot, message=FALSE, warning=FALSE}
excellent = brfss %>%
  filter(response == "Excellent") %>%
  group_by(year, locationabbr) %>% 
  summarise(mean_data_value = mean(data_value))

ggplot(excellent, aes(x = year, y = mean_data_value, color = locationabbr)) +
  geom_line(size = 0.5) +
  labs(
    x = "Year",
    y = "Average Data Value",
    title = "Average Data Value Across State Locations from 2002 - 2010"
    ) +
  theme(legend.title = element_blank())
```

  * Make a two-panel plot showing, for the years 2006, and 2010, distribution of data_value for responses (“Poor” to “Excellent”) among locations in NY State

The two-panel plot below shows the distribution of `data_value` for responses "Poor" to "Excellent" among locations in NY state for 2006 and 2010 side by side.

```{r two panel plot, message=FALSE, warning=FALSE}
responses = brfss %>% 
  filter(locationabbr %in% "NY",
         year %in% c(2006, 2010)) %>% 
  ggplot(aes(x = response, y = data_value, fill = response)) +
  geom_col(alpha = .5) +
  geom_smooth(se = FALSE) + 
  facet_grid(. ~ year) + 
  labs(x = "Response",
       y = "Data Value",
       title = "Distribution of Data Value for Responses in NY State in 2006 vs 2010"
  )

responses
```




# PROBLEM 3

  * Load, tidy, and otherwise wrangle the data. Your final dataset should include all originally observed variables and values; have useful variable names; include a weekday vs weekend variable; and encode data with reasonable variable classes. Describe the resulting dataset (e.g. what variables exist, how many observations, etc).


```{r cleaning accel, message=FALSE, warning=FALSE}
accel = read_csv(file = "./data/accel_data.csv") %>% 
  janitor::clean_names() %>% 
  pivot_longer(
    activity_1:activity_1440,
    names_to = "minute",
    values_to = "activity_counts") %>% 
  mutate(
    type = if_else(
      day %in% c("Saturday", "Sunday"),"weekend", "weekday"),
    day = as.factor(day), #since we can consider it to be ordered 
    day_id = as.numeric(day_id),
    minute = str_replace(minute, "activity_", ""),
    minute = as.numeric(minute),
    minute = as.numeric(minute),
    activity_counts = as.numeric(activity_counts))

head(accel)
```

#DESCRIBE RESULTING DATASET


  * Traditional analyses of accelerometer data focus on the total activity over the day. Using your tidied dataset, aggregate accross minutes to create a total activity variable for each day, and create a table showing these totals. Are any trends apparent?

```{r total counts, message=FALSE, warning=FALSE}
accel_total = accel %>% 
  group_by(week, day) %>% 
  summarize(total_activity_counts = sum(activity_counts))

kable(accel_total)
```

#ARE THERE ANY TRENDS?

  * Accelerometer data allows the inspection activity over the course of the day. Make a single-panel plot that shows the 24-hour activity time courses for each day and use color to indicate day of the week. Describe in words any patterns or conclusions you can make based on this graph
  
```{r single panel plot, message=FALSE, warning=FALSE}
accel_plot = accel %>% 
  mutate(hour = minute %/% 60) %>% #the plot is easier to read with an hour variable
  group_by(day, hour) %>%
  summarize(total_activity_counts = sum(activity_counts)) %>%
  ggplot(aes(x = hour, y = total_activity_counts, color = day)) + 
    geom_line() + 
  scale_x_continuous(breaks = 0:24) +
  scale_y_continuous(breaks = c(0, 50000, 100000, 150000, 200000, 250000)) +
  labs(
    x = "Hour",
    y = "Total Activity Counts (in thousands)", 
    title = "24-hour Activity Time Courses Per Day"
  )

accel_plot
```
