---
title: "HW3"
output: github_document
---

```{r setup, message=FALSE}
library(tidyverse)
library(readxl)
```

```{r loading data}
library(p8105.datasets)
data("instacart")
```

[Insert short description including: ]

  * size and structure
  * describing some key variables
  * illustrative examples of observations

[Insert answers to questions: ]

  * how many aisles, which aisles are most items ordered from
  * plot showing number of items ordered in each aisle, limiting this to aisles with more than 10000 items ordered. Arrange aisles sensibly, and organize your plot so others can read it.

```{r aisles cleaning and plot, message=FALSE}
instacart %>%
  group_by(aisle) %>% 
  tally() %>% #creates n variable so that I can make aisle a factor variable with levels based on n
  ungroup() %>% 
  mutate(aisle = factor(aisle, levels = aisle[order(n)])) %>% #ordered to make plot from most to least items ordered
  filter(n > 10000) %>% 
  ggplot(aes(x = aisle, y = n)) +
  geom_col() +
  coord_flip() +
  labs(title = "Items Ordered per Aisle",
       x = "Aisle",
       y = "Count of Items Ordered"
       ) +
  theme(text = element_text(size = 10))
```

  * Make a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”. Include the number of times each item is ordered in your table.



  * table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format this table for human readers (i.e. produce a 2 x 7 table)